Dynamic Modeling, Parameter Estimation and Uncertainty Analysis in \proglang{R}

### 2016-10-02 article.tex
theory ode model

\subsection{Ordinary differential equation models}

>>>>

\subsection{Ordinary differential equation models}
Object of the dynamic modeling are systems with states $x$, usually quantifying the involved species, that evolve over time as e.g. given by chemical reaction networks. The time evolution of the states is expressed via a set of ODEs, $\dot x = f(x)$. 
Typical examples follow the law of mass action:
\begin{itemize}
	\item $\dot x_A = k_p$ constant production of $x_A$ ($\emptyset \rightarrow A$)
	\item $\dot x_C = k_c \cdot x_A x_B = - \dot x_A = - \dot x_B$ complex formation ($A+B \rightarrow C$)
	\item $\dot x_C = - k_d x_C$ constant degradation of $x_C$ ($C \rightarrow \emptyset$)
\end{itemize}
generally written as $\dot x_l = \sum_{j = 1\dots m}  k_{j} \prod_{i = 1\dots n} \nu_{i,j} x_i^{\nu_{i,j}}$ for a system of $n$ states and $m$ reactions with rate constants $p = \{ k_j, j = 1\dots m\}$ and stoichiometric constants $\nu_{i,j} \in \mathbb{Z}$. 
An even more general expression of the system is given by $\dot x = S v(x,p)$ with the stoichiometry matrix $S$ containing the stoichiometric constants $\nu_{i,j}$ and the flux vector $v(x,p)$.
This expression also holds for more general systems that do not follow mass action kinetics. Moreover the system may be explicitly time-dependent and it may contain forcings $u(t)$ which e.g. describe the external stimulation of the system with a compound, resulting in the general form: \begin{align*}\dot x = S \cdot v(x,p,u,t).\end{align*}

The aim of parameter estimation will be the determination of the free parameters of the system. Besides the rate constants, they include parameters of the forcings $u(t)$ and initial values of the states at time point zero. It is generally assumed that the stoichiometry of the system is known.
